---
title: "Milestone Report"
author: "Michael Berger"
date: "2 December 2018"
output: 
  html_document: 
    fig_height: 7
    fig_width: 11
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(include = FALSE)
knitr::opts_knit$set(root.dir = '~/Studies/Coursera/10-Capstone/corpus/en_US/')
#setwd('~/Studies/Coursera/10-Capstone/corpus/en_US/')
 library(knitr)
# library(readtext)
 library(quanteda)
# library(SnowballC)
# library(ggplot2)
 library(LaF)
# library(plyr)
# library(dplyr)
# library(tidyr)
# library(plotly)
 library(data.table)
# library(stringi)
# library(hunspell)
# library(gridExtra)
# library(viridis)
 library(RColorBrewer)
```

##Abstract
In this project we will try to develop a word prediction system at typing. For this task we are provided by the JHU partner [SwiftKey](https://www.microsoft.com/en-us/swiftkey) with a large corpus of texts. We are supposed to use these texts to extract some frequent combinations of words we use in daily life to build our prediction algorithm from.  
 
## Summaries
First, let's see the scale of the problem. Since the files are very large and R is only able to work in-memory, I will use operating system's wc operation to count lines, words and characters.
    
```{r summaries, echo=FALSE, cache=T}

wc <- function(path)
{
    size <- round(file.size(path)/(2^20))
    path = paste("'",path, "'", sep = "")
    w1 <- system(paste("wc -lwc",path), intern=TRUE)
    
    wsplit <- unlist(strsplit(w1," +"))
    wnum <-as.numeric(wsplit[2:4])
    return(data.frame('file' = wsplit[5], 'size,MiB' = size, 'lines'=wnum[1], 'words'=wnum[2], 'chars'=wnum[3], stringsAsFactors = F))
}

files = list.files(".",pattern = "en*", full.names = F)

files = lapply(files, FUN = wc)# determine_nlines)

#filesbk <- files #files <- filesbk  , stringsAsFactors = F

files <- data.frame(ldply(files, rbind))
rownames(files) = files$file

files[nrow(files)+1,2:5] <- apply(files[,2:5],2,sum)
files[nrow(files),1] <- "Total"

files$wordsPerLine <- round(files$words/files$lines)
files$charsPerLine <- round(files$chars/files$lines)

kable(files, row.names = F,format.args = list(decimal.mark = ".", big.mark = ","))
```
 
As one can see the numbers are truly enormous. I will try to use a 1% of the lines to perform some preliminary analyses.
Also, looking ahead, running the analysis on all three files yelded pretty much similar results, so I will present it for only 'en_US.blogs.txt' file to avoid clogging up the report with duplicate information.

##Creating corpus
```{r oneprcent , echo = F}
oneprc <- round(files$lines * .1, -3) 
names(oneprc)<- files$file
kable(oneprc, col.names = "1%")
```  

For each file 1% of the lines is roughly 100,000 lines, a sizeable sample, but not too big, allowing to execute operations in a matter of seconds and without heavy load on memory. 
So in this step I:  
-Read 100,000 random lines from a file  
-Convert everything to lower case and remove foreign languages (non-ASCII) characters  
-Create corpus object  
 
```{r Corpus, comment=NA, cache=T }
currentFile = 'en_US.blogs.txt'
if (is.na(get_lines(currentFile,1:2)[1])) # check that I can read from file
  stop("couldn't read from file. probably path is wrong")
set.seed(42)
Lines <- sample_lines(currentFile, 100000, files[currentFile,]$lines) #sample some lines from a file
Encoding(Lines) <- "latin1"  #remove non-ascii chars 
Lines <- iconv(Lines, "latin1", "ASCII", sub="")
Corp <- corpus(char_tolower(Lines)) # everything to lowerCase and create corpus
paste(head(texts(Corp)))
```

##Tokenization
Creating word tokens, while removing unwanted tokens:  
  - numbers  
  - punctuation  
  - symbols  
  - separators  
  - hyphens  
  - URLs  
  - twitter tags (but leaving the word after # in #hashtag)  

```{r Tokens, cache=T}
Toks<- tokens(Corp, what = "word", remove_numbers = T, remove_punct = T, remove_symbols = T,
  remove_separators = T, remove_twitter = T, remove_hyphens = T, remove_url = T)
```  

And:  
  - twitter rt tag  
  - english stopwords  
  - profane words  
  
```{r stopwords_and_profane, cache=T}
l <- determine_nlines('profanity_words.txt') #read profanity vocabulary and convert to lower
profane <-char_tolower(get_lines('profanity_words.txt', 1:l))
profane <- profane[!is.na(profane)]
Toks<- tokens_remove(Toks, c(stopwords("english"), profane, "rt"))  # also remove ReTweet tag
```    
#Stemming  
Stemming is an operation of converting word to its root form. Example: caresses -> caress.  
I decided not to do stemming because:  
  - it converts some words in a strange manner, like: ponies -> poni.  
  - the model will output stemmed words as prediction which is undesirable.  
    
##Document Frequency Matrix  
A DFM is a matrix that stores number of appearances of each token. Essentially, it extracts all the unique words from a text together with total number of appearances of this word. At this step DFM is useful to spot unwanted words. 
  
```{r 1-gram_DFM, cache=T}
Dfm1gr <- dfm(Toks, tolower = TRUE, stem = F) 
```
  
###Check that non-ASCII characters actually got removed  
  If present, these chars should be immediately seen at the beginning or at the end of sorted tokens list.  
  
```{r nonASCII_check ,echo = F}
features <- Dfm1gr@Dimnames$features
lvls <- features[order(features)]
cat(head(lvls,50), sep ="; ")
cat(tail(lvls,50), sep ="; ")
```
And I see that there are none. Even though there are some pseudo-words here, like "100,00th", these will be filtered out at a later step as having too low a frequency in the n-grams.

##Statistics
Since the data is very big, I have to find ways to make it smaller. 
First of all, it will become smaller due to structure of the model which compressees the data to just a database of unique tokens with their frequencies. 
Another way is to filter out terms that are of very low frequency.  
For this I need to determine how much data can be filtered out and still cover 90% of the words.

  
```{r coverage_graph, cache=T}
freq1gr <- colSums(Dfm1gr)
freq1gr <- data.frame(words = names(freq1gr), freq = freq1gr , stringsAsFactors = F)
freq1gr  <- freq1gr[order(freq1gr$freq, decreasing = T),]
freq1gr$wc <- 1:nrow(freq1gr)
freq1gr$cs <- cumsum(freq1gr$freq)
freq1gr$perc <- freq1gr$cs/tail(freq1gr$cs,1)

closest <- function(w,x) #find closest value to x in a vector w. w must be sorted.
{
  dt = data.table(w, val = w) 
  setattr(dt, "sorted", "w")
  return(dt[J(x), .I, roll = "nearest", by = .EACHI])
}

cl<-closest(freq1gr$perc,c(.5, .9, .99))
cl$p <- round((cl$I/nrow(freq1gr))*100)
cl$col <- brewer.pal(3, "Set1")
coverage <- ggplot(data = freq1gr ,aes(x = wc, y = perc)) + geom_line() +
              geom_hline(aes(yintercept =  w, col = col),data = cl, show.legend = F)  +
              geom_label(aes(x = I, y = w+.04, label=paste(I, " ≈ ", p ,"%"), col = col ), data = cl, show.legend = F) + xlab("num of unique words") + ylab("% most frequent tokens") 
coverage
```
  
So I see that the coverage graph grows very fast. About 20% most frequent tokens cover 90% of all unique words in a 100k lines sample.  
  
##N-grams
N-grams are a sequences of n words taken from a text and incorporated as a single token.  
So if a text was "The quick brown fox jumps over  another quick brown fox" then 3-grams for it would be:

```{r brown_fox, echo=FALSE}
fox = "The quick brown fox jumps over another quick brown fox"
foxtok = tokens(fox,ngrams = 3, remove_number =T)
myfun <- function(x) {paste("\n", strrep(" ",cumsum(nchar(x)+1)))}
lal <- sapply(strsplit(fox," "),myfun)
cat(unlist(foxtok), sep = lal)
```
  
The next step is to create Document-feature matrix (DFM) out of these 3-grams. So for example in the case of 3-grams each token represents 3 consecutive words. So all the tokens in our example will have 1 appearance except for "quick_brown_fox", which will have 2.

I will create 1, 2 and 3-grams for the document to get the general picture of the data. 1-gram is just dfm of single words which I've already created earlier. 
```{r 2_3_grams, cache=T}
bigrams <- tokens_ngrams(Toks, 2)
trigrams <- tokens_ngrams(Toks, 3)

Dfm2gr <- dfm(bigrams, tolower = TRUE, stem = F)
Dfm3gr <- dfm(trigrams, tolower = TRUE, stem = F)
```

  
A plot of most frequent 1,2,3-grams shows us that even such simple algorithm is indeed able to locate frequent combinations of words we use in daily life.  
  
  
```{r plot_ngrams, cache=T}

tf <- topfeatures(Dfm2gr,30, scheme =  "count")
tf2 <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)

tf <- topfeatures(Dfm3gr,30, scheme =  "count")
tf3 <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)
                   
plt1grm <- ggplot(head(freq1gr,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') + xlab("") +
        coord_flip() + ylab("Number of Appearances") + ggtitle("most frequent 1grams") 

plt2grm <- ggplot(head(tf2,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') + xlab("") +
        coord_flip() + ylab("Number of Appearances") + ggtitle("most frequent 2grams") 
        
plt3grm <- ggplot(head(tf3,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') + xlab("") +
        coord_flip() + ylab("Number of Appearances") + ggtitle("most frequent 3grams") 
      
grid.arrange(plt1grm, plt2grm, plt3grm, ncol=3)

```


```{r save_ngrams, eval=FALSE, include=FALSE}
save(Dfm1gr,Dfm2gr,Dfm3gr, file = "123grams.rData")
```


## Next steps

So judging from this overview, a straightforward solution to the project's goal seems simple:  
-Pre-calculate a frequencies database for all 1,2,3,4, maybe 5-grams.  
-For every new word in a sequence that a user types, find in the database all n-grams that begin similarly  
-Calculate a degree of fit for every found n-gram according to some formula  
-Select n-gram with best fit.  

## Implementation

Since the given data is truly enormous and can't fit into memory, I will have to process it in chunks, saving only ngrams and frequencies in an SQL database.

## Conclusion 
In this report I have performed a brief overview of the data that is avilable and arrived at an idea as to how it may be useful for creating a prediction model. 








