---
title: "Milestone Report"
author: "Michael Berger"
date: "2 December 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/Studies/Coursera/10 - Capstone/corpus/en_US/')
library(knitr)
library(readtext)
library(quanteda)
library(SnowballC)
library(ggplot2)
library(LaF)
library(dplyr)
library(tidyr)
library(plotly)
library(data.table)
library(stringi)
library(hunspell)
library(gridExtra)
```
 
 
  summaries:
    number of lines 
    num words + per line + distribution  (ntokens, ntype)
    
```{r summaries, echo=FALSE, cache=T}

folder = "/home/michael/Studies/Coursera/10 - Capstone/corpus/en_US"

files = list.files(folder, full.names = F)

lines = sapply(files, FUN = determine_nlines)

files <- data.frame(numlines = lines, stringsAsFactors = F)#, lines = as.vector(t(lns), mode = "any")

kable(files)
```
 

 
 
  Read a limited random sample of text file
  Everything to lower
  remove foreign languages characters (non-ASCII characters)
  
  create corpus
 
```{r Corpus}
 
if (is.na(get_lines('en_US.twitter.txt',1:2)[1])) # check that I can read from file
  stop("shnitzl: couldn't read from file. probably path is wrong")

set.seed(42)
currentFile = 'en_US.twitter.txt'
twitLines <- sample_lines(currentFile, 50000, files[currentFile,]) #sample some lines from a file

Encoding(twitLines) <- "latin1"  #remove non-ascii chars 
twitLines <- iconv(twitLines, "latin1", "ASCII", sub="")

twitCorp <- corpus(char_tolower(twitLines)) # everything to lowerCase and create corpus

head(texts(twitCorp))

```

 read profanity vocabulary and convert to lower

```{r profane}
profane <-char_tolower(get_lines('profanity_words.txt', 1:450))
profane <- profane[!is.na(profane)]
tail(profane)
```


  
  create word tokens, while removing unwanted tokens:
    v numbers 
    v punctuation 
    v symbols
    v separators
    v hyphens
    v url    
    v twitter (leave twe word after # in #hashtag)
    v twitter rt tag (tokens_remove)   
    v english stopwords
    v profane words
    
```{r}
twitToks<- tokens(twitCorp, what = "word", remove_numbers = T, remove_punct = T,
  remove_symbols = T, remove_separators = T,
  remove_twitter = T, remove_hyphens = T, remove_url = T)
twitToks<- tokens_remove(twitToks, c(stopwords("english"), profane, "rt"))  # also remove ReTweet tag
```
    
    
  I decided not to do stemming because 
    1) it converts some words in a strange manner 
    2) the model will output stemmed words as prediction.
    
  
  create dfm  

  
  
```{r}

Dfm1gr <- dfm(twitToks, tolower = TRUE, stem = F) # no stemming required i think
rs<-colSums(Dfm1gr) # maybe its betteer to use docfreq()
rs <- data.frame(word = names(rs), freq = rs)
tf <- rs[order(rs$freq, decreasing = T ),]

#tf <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)

Plot <- ggplot(head(tf,30), aes(x=reorder(word, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') +
        coord_flip() + xlab("frequency") + ylab("Number of Words") + ggtitle("1-grams most frequent tokens")
Plot
```
  
  check non-ASCII characters actually got removed  
  
```{r echo=FALSE}
# with this you can check if there are some non-ascii chars remaining in dfm like chinese or russian 
lvls <- levels(rs$word)
n = 1
#lvls[n:(n+100)]
cat(head(lvls,100), sep ="; ")
```

  Create ngrams: 1,2,3


```{r}
bigrams <- tokens_ngrams(twitToks, 2)
trigrams <- tokens_ngrams(twitToks, 3)
# (head(bigrams))
# (head(trigrams))

Dfm2gr <- dfm(bigrams, tolower = TRUE, stem = F)
Dfm3gr <- dfm(trigrams, tolower = TRUE, stem = F)
```

  
  barplot with 30 most frequent 1,2,3 grams




  determine how much data i need to cover 50% and 90% of the words
  plot a graph of coverage
  
  
```{r}
freq1gr <- docfreq(Dfm1gr)

freq1gr <- data.frame(words = names(freq1gr), freq = freq1gr , stringsAsFactors = F)

freq1gr  <- freq1gr[order(freq1gr$freq, decreasing = T ),]

freq1gr$wc <- 1:nrow(freq1gr)

freq1gr$cs <- cumsum(freq1gr$freq)

freq1gr$perc <- freq1gr$cs/tail(freq1gr$cs,1)


closest <- function(w,x)
{
  dt = data.table(w, val = w) 
  setattr(dt, "sorted", "w")
  return(dt[J(x), .I, roll = "nearest", by = .EACHI])
}

cl <-rbind( closest(freq1gr$perc,.5),closest(freq1gr$perc,.9),closest(freq1gr$perc,.99))
cl$p <- floor(cl$I/nrow(freq1gr)*100)

cumulative <- ggplot(data = freq1gr ,aes(x = wc, y = perc)) + geom_line() +
              geom_hline(aes(yintercept =  w, col = c("blue","red", "green")),data = cl, show.legend = F)  +
              geom_label(aes(x = I, y = w+.04, label=paste(I, " ≈ ", p ,"%"), col = c("blue","red", "green")), data = cl, show.legend = F)

cumulative
                     
                     
                     
plt1grm <- ggplot(head(freq1gr,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') +
        coord_flip() + xlab("frequency") + ylab("Number of Words") + ggtitle("1-grams most frequent tokens") 
      

plt1grm


```
  
  27% most frequent tokens cover 90% of all unique words in 50k lines sample. 
  
  
```{r}
tf <- topfeatures(Dfm2gr,30, scheme =  "docfreq")
tf2 <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)

tf <- topfeatures(Dfm3gr,30, scheme =  "docfreq")
tf3 <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)

plt2grm <- ggplot(head(tf2,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') +
        coord_flip() + xlab("frequency") + ylab("Number of Words") + ggtitle("2-grams most frequent tokens") 
        
plt3grm <- ggplot(head(tf3,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') +
        coord_flip() + xlab("frequency") + ylab("Number of Words") + ggtitle("3-grams most frequent tokens") 

  
grid.arrange(plt2grm, plt3grm, ncol=2)

```











