---
title: "Milestone Report"
author: "Michael Berger"
date: "2 December 2018"
output: 
  html_document: 
    fig_height: 7
    fig_width: 11
---

```{r first, include=FALSE}
#knitr::opts_chunk$set(include = FALSE)
knitr::opts_knit$set(root.dir = '~/Studies/Coursera/10-Capstone/corpus/en_US/')
setwd('~/Studies/Coursera/10-Capstone/corpus/en_US/')
library(knitr)
library(readtext)
library(quanteda)
library(SnowballC)
library(ggplot2)
library(LaF)
library(plyr)
library(dplyr)
library(tidyr)
library(plotly)
library(data.table)
library(stringi)
library(hunspell)
library(gridExtra)
library(viridis)
library(RColorBrewer)
```


##Abstract
In this project we will try to develop next word prediction system at typing. For this task we are provided by the JHU partner swiiftkey with data  
The data provided  
 
## Summaries
First, let's see the scale of the problem.  
Since the files are very large and R is only able to work in-memory, I will use operating system's wc operation to count lines, words and characters.
    
```{r summaries, echo=FALSE, cache=T}

wc <- function(path)
{
    size <- round(file.size(path)/(2^20))
    path = paste("'",path, "'", sep = "")
    w1 <- system(paste("wc -lwc",path), intern=TRUE)
    
    wsplit <- unlist(strsplit(w1," +"))
    wnum <-as.numeric(wsplit[2:4])
    return(data.frame('file' = wsplit[5], 'size,MiB' = size, 'lines'=wnum[1], 'words'=wnum[2], 'chars'=wnum[3], stringsAsFactors = F))
}

files = list.files(".",pattern = "en*", full.names = F)

files = lapply(files, FUN = wc)# determine_nlines)

#filesbk <- files #files <- filesbk  , stringsAsFactors = F

files <- data.frame(ldply(files, rbind))
rownames(files) = files$file

files[nrow(files)+1,2:5] <- apply(files[,2:5],2,sum)
files[nrow(files),1] <- "Total"

files$wordsPerLine <- round(files$words/files$lines)
files$charsPerLine <- round(files$chars/files$lines)

kable(files, row.names = F,format.args = list(decimal.mark = ".", big.mark = ","))
```
 
As one can see the numbers are truly enormous. I will try to use a 1% of the lines to perform some preliminary analyses.


##Creating corpus
  Read a limited random sample of text file   
```{r}
oneprc <- round(files$lines * .1, -3) 
names(oneprc)<- files$file

kable(oneprc, col.names = "1%")
```  
  A 100,000 lines seems to be a reasonable sample. Also, I will perform analysis on one of the files, since text is always text.
  Then convert everything to lowerCase and remove foreign languages characters (non-ASCII characters)
  
  create corpus
 
```{r Corpus}
 

if (is.na(get_lines('en_US.twitter.txt',1:2)[1])) # check that I can read from file
  stop("couldn't read from file. probably path is wrong")

set.seed(42)
currentFile = 'en_US.twitter.txt'
Lines <- sample_lines(currentFile, 100000, files[currentFile,]$lines) #sample some lines from a file

Encoding(Lines) <- "latin1"  #remove non-ascii chars 
Lines <- iconv(Lines, "latin1", "ASCII", sub="")

Corp <- corpus(char_tolower(Lines)) # everything to lowerCase and create corpus

paste(head(texts(Corp)))

```

 read profanity vocabulary and convert to lower

```{r profane}
l <- determine_nlines('profanity_words.txt')
profane <-char_tolower(get_lines('profanity_words.txt', 1:l))
profane <- profane[!is.na(profane)]
```


  
##Creating word tokens, while removing unwanted tokens:  
    - numbers  
    - punctuation  
    - symbols  
    - separators  
    - hyphens  
    - url 
    - twitter (leaving the word after # in #hashtag)  
    - twitter rt tag
    - english stopwords  
    - profane words  
    
```{r}
Toks<- tokens(Corp, what = "word", remove_numbers = T, remove_punct = T,
  remove_symbols = T, remove_separators = T,
  remove_twitter = T, remove_hyphens = T, remove_url = T)
Toks<- tokens_remove(Toks, c(stopwords("english"), profane, "rt"))  # also remove ReTweet tag
```
    
    
I decided not to do stemming because:  
  - it converts some words in a strange manner  
  - the model will output stemmed words as prediction which is undesired behaviour   
    
##Document Frequency Matrix  

  
```{r}

Dfm1gr <- dfm(Toks, tolower = TRUE, stem = F) # no stemming required i think


#tf <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)


```
  
###Check that non-ASCII characters actually got removed  
  If present, these chars should be immediately seen at the beginning or at the end of sorted tokens list.
  
```{r echo=FALSE}
# with this you can check if there are some non-ascii chars remaining in dfm like chinese or russian 
rs <-colSums(Dfm1gr) # maybe its betteer to use docfreq()
rs <- data.frame(word = names(rs), freq = rs)
tf <- rs[order(rs$freq, decreasing = T ),]
lvls <- levels(rs$word)
n = 1
#lvls[n:(n+100)]
cat(head(lvls,50), sep ="; ")
cat(tail(lvls,50), sep ="; ")
```
And I see that there are none. Even though there are some pseudo-words here, like "100,00th", these will be filtered out as having too low a frequency in the Document Feature Matrix.

##Statistics
Since the data is very big, I have to find ways to make it smaller. 
Intuition tells me that 
One way is to filter out terms that are of very low frequency. 
For this I need to determine how much data I need to cover 50% and 90% of the words

  
```{r}
freq1gr <- colSums(Dfm1gr)
freq1gr <- data.frame(words = names(freq1gr), freq = freq1gr , stringsAsFactors = F)
freq1gr  <- freq1gr[order(freq1gr$freq, decreasing = T ),]
freq1gr$wc <- 1:nrow(freq1gr)
freq1gr$cs <- cumsum(freq1gr$freq)
freq1gr$perc <- freq1gr$cs/tail(freq1gr$cs,1)

closest <- function(w,x)
{
  dt = data.table(w, val = w) 
  setattr(dt, "sorted", "w")
  return(dt[J(x), .I, roll = "nearest", by = .EACHI])
}

#cl <-rbind( closest(freq1gr$perc,.5),closest(freq1gr$perc,.9),closest(freq1gr$perc,.99))

cl<-closest(freq1gr$perc,c(.5, .9, .99))

cl$p <- round((cl$I/nrow(freq1gr))*100)

cl$col <- brewer.pal(3, "Set1")

coverage <- ggplot(data = freq1gr ,aes(x = wc, y = perc)) + geom_line() +
              geom_hline(aes(yintercept =  w, col = col),data = cl, show.legend = F)  +
              geom_label(aes(x = I, y = w+.04, label=paste(I, " ≈ ", p ,"%"), col = col ), data = cl, show.legend = F) + xlab("num of unique words") + ylab("% most frequent tokens") #+ brewer.pal(7, "BrBG")  #scale_fill_brewer(palette="Dark2")

coverage
```
  
So I see that the coverage graph grows very fast which means that about 20% most frequent tokens cover 90% of all unique words in a 100k lines sample. 
  
  
##N-grams
N-grams are a sequences of n words taken from a corpus and incorporated as a single token.
So if a corpus was "The quick brown fox jumps over the lazy dog" then 3-grams for it would be:

```{r echo=FALSE}
fox = "The quick brown fox jumps over the lazy dog "
dfoxm = tokens(fox,ngrams = 3, remove_number =T)
myfun <- function(x) {paste("\n", strrep(" ",cumsum(nchar(x)+1)))}
lal <- sapply(strsplit(fox," "),myfun)
cat(unlist(dfoxm), sep = lal)
```
  
  
```{r}
bigrams <- tokens_ngrams(Toks, 2)
trigrams <- tokens_ngrams(Toks, 3)

Dfm2gr <- dfm(bigrams, tolower = TRUE, stem = F)
Dfm3gr <- dfm(trigrams, tolower = TRUE, stem = F)
```

  
A plot of most frequent 1,2,3-grams shows us that even such simple algorithm is indeed able to locate most frequent combinations of words we use in daily life.  
  
  
```{r}
knitr::opts_chunk$set(fig.width = 9)
tf <- topfeatures(Dfm2gr,30, scheme =  "count")
tf2 <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)

tf <- topfeatures(Dfm3gr,30, scheme =  "count")
tf3 <- data.frame(words = names(tf), freq = tf , stringsAsFactors = F)


                     
plt1grm <- ggplot(head(freq1gr,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') + xlab("") +
        coord_flip() + ylab("Number of Appearances") + ggtitle("most frequent 1grams") 

plt2grm <- ggplot(head(tf2,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') + xlab("") +
        coord_flip() + ylab("Number of Appearances") + ggtitle("most frequent 2grams") 
        
plt3grm <- ggplot(head(tf3,30), aes(x=reorder(words, freq), y=freq)) + 
        geom_bar(stat="Identity", fill='red') + xlab("") +
        coord_flip() + ylab("Number of Appearances") + ggtitle("most frequent 3grams") 
      

  
grid.arrange(plt1grm, plt2grm, plt3grm, ncol=3)

```











